Comparing vector-based and ACT-R memory models of tag retrieval:\protect\\User-customized hashtag/tag prediction on Twitter and StackOverflow

### 2014-07-14 writeups/dissertation/dissertation.tex
foo

% If the toc is too detailed, try limiting the depth of printed sections in the TOC:
%\setcounter{tocdepth}{2}

@@

\subsubsection{StackOverflow Dataset}

@@

\subsubsection{Twitter Dataset}

@@

\subsection{Pre-processing the Text}

@@

\subsection{Stop Words and Entropy Weighting}

@@

\subsection{Attenuating Twitter Spam Tweets}

Although Twitter has an infrastructure to easily report and remove spam accounts, the spam that does get through can contain tweets with numerous hashtags.
For example, some spam accounts generate tweets where all words in the tweet are hashtags, or generate multiple tweets with repeating content and hashtags.
Feeding these spam tweets into statistical models that generate hashtag predictions can bias the models' chosen hashtags.

\textcite{Mazzia2009} developed a novel approach to dealing with hashtag spam on Twitter.
When training his model, it attenuates the importance of multiple tweets from the same user that contain the same hashtags.
As a user generates more than 10 tweets that contain the same hashtag, all tweets afterwards that contain this hashtag will be attenuated so that the overall weight is no more than that for 10 tweets.
When testing the model, only at most 1 post from a user is included in the test set, since it is much easier to suggest relevant hashtags to a spammer than it is for an actual user.
This exact technique may not be used for this analysis, but certainly some form of attenuating or identifying Twitter spam tweets will be explored.

\subsection{Experiments}

Each proposed experiment aims to answer a specific research question in this study.
The results from Experiment 1 will show how a user's prior hashtag use influences their likelihood of using that hashtag in the future, and test whether the observed data fit the ACT-R frequency/recency decay rate theory.
The output from Experiment 1 will be a model of how a user's prior hashtag use influences the probability of using that hashtag in the future for both the Twitter and StackOverflow datasets.
Experiment 2 will use the results from Experiment 1 and develop the best method to incorporate user's past tagging history in the random-permutations model, where user history is currently not being used.
Experiment 3 will explore various methods for incorporating word order information into the ACT-R Bayesian model, where this information is currently not being used.
The outputs of Experiment 2 and 3 are improved ACT-R and random-permutation models that have properly incorporated the strengths of each other: word-order strength from random permutations and user history from ACT-R.
Finally, Experiments 4 and 5 will directly compare model accuracy between the two models and evaluate overall performance.
Experiment 4 is focused primarily on hashtag prediction when no user prior history is available, while Experiment 5 will test how much performance improves when that history is available.

\subsubsection{Preparation Experiment: StackOverflow Dataset Modifications}

\textcite{Stanley2013} developed a model for the StackOverflow dataset that was based on all posts up to April, 2011.
This model will be used for this research, but the dataset will be updated to the most recent release.
When moving from the older dataset to the newest dataset, several performance assessments will be made.

First, model performance will be compared between the newest dataset and the Apr' 11 dataset to ensure that performance is consistent.
This is also a way to verify that the newest dataset was correctly pre-processed and loaded.

Next, incremental improvements will be made to the pre-processing stages and performance will be compared after each change.
For example, in \textcite{Stanley2013} the technique used to chunk the words does not ensure that tags appearing in the post remain as a single word unit.
So a word like \emph{C\#} in a post was chunked as two separate words (\emph{C} and \emph{\#}).
This reduces predictive power of the model since it muddles the highly-predictive co-occurrence observations where a tag appears in the post.
So the chunking technique will be changed to take into account tags in the post, and its effect on performance will be evaluated.

\subsubsection{Experiment 1: Analyze Recency for User-Generated Hashtags}

@@

\subsubsection{Experiment 2: Incorporating Recency Into Random Permutations}

@@

\subsubsection{Experiment 3: Incorporating Word Order Into ACT-R}

How can word-order information be incorporated into the ACT-R Bayesian model?
Currently ACT-R uses a word co-occurrence matrix that does not include word position.
A co-occurrence is counted if a word and a tag co-occur anywhere in a tweet or post.
However, for Twitter in particular, it may be the case that the few words immediately preceding hashtags are highly predictive, and the word position for those words should be included in the model.
So a method to include word order into a Bayesian model will be developed.

One method to include word order in a Bayesian co-occurrence matrix is to represent word position on a third, orthogonal dimension, and turn the matrix into a 3-dimensional array.
So a co-occurrence is represented by a triplet: context word by word position relative to tag by tag.
When a retrieval request for a tag is made, the strength of association between each word in context (and its position) and the tags in the co-occurrence array would be computed.
That computation could be made tractable by storing only a small set of values on the order axis and binning over that dimension.

For example, one could store only three word order values: near, medium, far, which is equivalent to storing three separate co-occurrence matrices.
So when a retrieval request is made (e.g., when a user writes the \# character when composing a tweet), the strength of association for each word would be computed using the appropriate matrix.
That is, the ``far'' matrix would be used for words far away from the cursor position, ``near'' for words just written, and ``medium'' for words mid distance from the current focus point.
This method seems straightforward to implement, but it is slightly unsatisfying since information is lost when binning is used.
It also does not scale, since an additional 2-dimensional sparse matrix must be added for every additional value stored on the word order dimension.
So I will try to discover a better technique to incorporate word order information into the ACT-R Bayesian model than the one just described, but will fall back to this technique if a better method is not found.

This updated model will then be evaluated on only the Twitter popular-hashtags dataset, since this dataset contains a large (yet not infinite) set of hashtags that occur in various positions throughout the tweet.
This model will not be evaluated on the StackOverflow dataset since chosen tags for a StackOverflow post are not embedded and mixed within a post, so word-order information is unavailable and most likely unused.

\subsubsection{Experiment 4: Top Hashtag Prediction}

After incorporating the relative strengths of each model into the other, the accuracy of the two models will be compared directly.
First both models will be evaluated on predicting tags based purely on context (i.e., no prior user history), using the popular-hashtags, popular-users, and StackOverflow datasets.
The results will show the performance differences for the contextual component of each model.

\subsubsection{Experiment 5: User-Customized Hashtag Prediction}

Next the full models (prior and context components) will be tested on the Twitter popular-users dataset and the StackOverflow dataset.
The popular-hashtags dataset will not be included in this analysis, since this subset was collected across the entire Twitter user population, so prior hashtag use is not available.
This test will directly compare performance differences for each model when generating user-customized tag predictions.
The main evaluation will be the amount of performance gain achieved when the models generate tags that are based on past user behavior compared to models that generate tags that are based purely on context.

\subsection{Summary}

>>>>

% If the toc is too detailed, try limiting the depth of printed sections in the TOC:
\setcounter{tocdepth}{2}

@@

\subsection{StackOverflow Dataset}

@@

\subsection{Twitter Dataset}

@@

\subsection{Aquiring Datasets}

\subsection{Tokenization}

@@

\subsection{Handling Stop Words}

@@

%FIXME: Add paragraph for freq weighting

\section{Results}

\subsection{Past User Behavior}

@@

\subsubsection{Method}

\subsubsection{Results}

\subsection{Adding Offset Term}

\subsubsection{Method}
\subsubsection{Results}

\subsection{Twitter Subsets}

\subsubsection{Method}
\subsubsection{Results}

\subsection{Frequency vs Entropy vs Stoplist Weighting}

\subsubsection{Method}
\subsubsection{Results}

\subsection{Dataset Size}

\subsubsection{Method}
\subsubsection{Results}

\subsection{Amount of Compression for Random Permutation}

\subsubsection{Method}
\subsubsection{Results}

\subsection{Method to Combine Prior and Context}

@@

\subsubsection{Method}
\subsubsection{Results}

\subsection{Word Order}

\subsubsection{Method}
\subsubsection{Results}

\subsection{Impact of Prior vs Context}

\subsubsection{Method}
\subsubsection{Results}

\subsection{User-customized Co-occurrence Matrix}

\subsubsection{Method}
\subsubsection{Results}

\section{General Discussion}

\section{Summary}
